---
title: "Plot performance summaries"
author: 
- name: Charlotte Soneson
- name: Angelo Duo
date: "`r Sys.Date()`"
package: DuoClustering2018
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Plot performance summaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  chunk_output_type: console
---

# Introduction

In this vignette we describe the basic usage of the `DuoClustering2018` package:
how to retrieve data sets and clustering results, and how to construct various
plots summarizing the performance of different methods across several data sets.

# Load the package

```{r}
suppressPackageStartupMessages({
  library(DuoClustering2018)
  library(plyr)
  library(ExperimentHub)
  library(SingleCellExperiment)
})
```

# Retrieve a data set

The clustering evaluation is based on 12 data sets (9 real and 3 simulated),
which are all provided via `ExperimentHub` and retrievable via this package. We
include the full data sets (after quality filtering of cells and removal of
genes with zero counts across all cells) as well as three filtered versions of
each data set (by expression, variability and dropout pattern, respectively),
each containing 10% of the genes in the full data sets.

To get an overview, we can list all records from this package that are available
in `ExperimentHub`:

```{r}
eh <- ExperimentHub()
query(eh, "DuoClustering2018")
```

The records with names starting in `sce_` represents (filtered or unfiltered)
data sets (in `SingleCellExperiment` format). The records with names starting in
`clustering_summary_` correspond to `data.frame` objects with clustering results
for each of the filtered data sets. Finally, the
`duo_clustering_all_parameter_settings` object contains the parameter settings
we used for all the clustering methods. For clustering summaries and parameter
settings, the version number (e.g., `_v1`) corresponds to the version of the
publication.

The records can be retrieved using their `ExperimentHub` ID, e.g.:

```{r}
eh[["EH1500"]]
```

Alternatively, the shortcut functions provided by this package can be used:

```{r}
sce_filteredExpr10_Koh()
```

# Read a set of clustering results

For each included data set, we have applied a range of clustering methods (see
the `run_clustering` vignette for more details on how this was done, and how to
apply additional methods). As mentioned above, the results of these clusterings
are also available from `ExperimentHub`, and can be loaded either by their
`ExperimentHub` ID or using the provided shortcut functions, as above. For
simplicity, the results of all methods for a given data set are combined into
one data frame. As an illustration, we load the clustering summaries for two
different data sets (`Koh` and `Zhengmix4eq`), each with two different gene
filterings (`Expr10` and `HVG10`):

```{r}
res <- plyr::rbind.fill(
  clustering_summary_filteredExpr10_Koh_v1(),
  clustering_summary_filteredHVG10_Koh_v1(),
  clustering_summary_filteredExpr10_Zhengmix4eq_v1(),
  clustering_summary_filteredHVG10_Zhengmix4eq_v1()
)
dim(res)
head(res)
```

# Define consistent method colors

For some of the plots generated below, the values will be colored according to
the clustering method. We can enforce a consistent set of colors for the methods
by defining a named vector of colors to use for all plots.

```{r}
method_colors <- c(CIDR = "#332288", FlowSOM = "#6699CC", PCAHC = "#88CCEE", 
                   PCAKmeans = "#44AA99", pcaReduce = "#117733",
                   RtsneKmeans = "#999933", Seurat = "#DDCC77", 
                   SC3svm = "#661100", SC3 = "#CC6677", 
                   TSCAN = "grey34",  ascend = "orange", SAFE = "black")
```

# Plot

Each plotting function described below return a list of `ggplot` objects. These
can be plotted directly, or further modified if desired.

## Performance

The `plot_performance()` function generates plots related to the performance of
the clustering methods. We quantify performance using the adjusted Rand Index
(ARI), comparing the obtained clustering to the true clusters. As we noted in
the publication, defining a true partitioning of the cells is difficult, since
they can often be grouped together in several different, but still
interpretable, ways. We refer to our paper for more information on how the true
clusters were defined for each of the data sets.

```{r}
perf <- plot_performance(res, method_colors = method_colors)
names(perf)
perf$median_ari_vs_k
perf$median_ari_heatmap_truek
```

## Stability

The `plot_stability()` function evaluates the stability of the clustering
results from each method, with respect to random starts. Each method was run
five times on each data set (for each k), and we quantify the stability by
comparing each pair of such runs using the adjusted Rand Index.

```{r}
stab <- plot_stability(res, method_colors = method_colors)
names(stab)
stab$stability_allmethods
stab$stability_heatmap_truek
```

## Entropy

In order to evaluate the tendency of the clustering methods to favor equally
sized clusters, we calculate the Shannon entropy of each clustering solution
(based on the proportions of cells in the different clusters) and plot this
using the `plot_entropy()` function. Since the maximal entropy that can be
obtained depends on the number of clusters, we use normalized entropies, defined
by dividing the observed entropy by `log2(k)`. We also compare the entropies for
the clusterings to the entropy of the true partition for each data set.

```{r}
entr <- plot_entropy(res, method_colors = method_colors)
names(entr)
entr$entropy_byds_byk
entr$normentropy_by_ds
```

## Timing

The `plot_timing()` function plots various aspects of the timing of the
different methods.

```{r}
timing <- plot_timing(res, method_colors = method_colors, scaleMethod = "RtsneKmeans")
names(timing)
timing$time_normalized_by_ref
```

## Differences in k

Most performance evaluations above are performed on the clustering solutions
obtained by imposing the true number of clusters. The `plot_k_diff()` function
evaluates the difference between the true number of cluster and the number of
clusters giving the best agreement with the true partition, as well as the
difference between the estimated and the true number of clusters, for the
methods that allow estimation of k.

```{r}
kdiff <- plot_k_diff(res, method_colors = method_colors)
names(kdiff)
kdiff$diff_kest_ktrue
```

# Session info

```{r}
sessionInfo()
```

